### 2014-03-27 paper/discussion.tex
adding bjoern comments
paper workin

Pipes included in objects can have additional effects beyond their designed intentions.

@@

\valkyrie{To discuss: printing media into tubes?  Design of interiors of objects more generally?  Need to cite that paper about balancing printed objects from SIGGRAPH.}

>>>>

The promise of 3D printing is that it allows for the manufacture of geometries that are impossible to create using traditional techniques like injection molding.  Pipes leverage that promise: designing the internal structure of an object was not previously feasible.  As printers at the hobbyist and corporate levels become more sophisticated and able to fabricate intricate geometries of ever more materials, we believe that our tool will still assist in design tasks.  Even if printers are capable of laying down conductive, optically transparent, or luminescent materials in-place, the process of designing their paths is still crucial.

Once objects have been designed and fabricated, the major flexibility for prototyping with pipes stems from the media inserted in them and the systems attached which sense changes in, or actuate, those media.  The five types of media--gas, liquid, solid, particulate, and threadable--give rise to unique possibilities in output (visual, aural, tactile/haptic, olfactory/gustatory) as well as input (touch/pressure and others).

\valkyrie{this section is probably my project for the weekend, or something.  I can't figure out how to improve it right now.  Tovi, we should talk about this in our meeting this afternoon.}

\tovi{There's lots of great ideas an examples in the inputs and outputs section, but the orgnization and coorespondence to the table is a little hard to follow. }

By crossing pipe types, topologies, and inserted media, we create a space of possible inputs and outputs that can be pipe-mediated.  We offer a table (see Figure \ref{fig:designspace}) containing these possibilities, referencing previous work on input/output techniques where appropriate, noting that many of these techniques have not been attempted using digital fabrication \bjoern{It may be a bit tricky to define what "using digital fab" means here. For example, Harrison 's pneumatically actuated buttons were done on a laser cutter, but also required manual gluing of other layers. Does that count or not?}.  We suggest new points in the design space that have been unexplored or not yet explored with digital fabrication.

\begin{figure*}[t]
\centering
    \includegraphics[width=\textwidth]{figures/designspace.pdf}
\caption{The design space of pipe-based interactions.  Existing systems are written in regular font.  Those created with fabrication are in {\color{blue}blue}. In \emph{{\color{red}red italic}} are unexplored interactions creatable with custom-fabricated pipes. \tovi{Im not sure how well this table will work. Division between input and output should be more prominent.} \valkyrie{one possibility is making the table larger and adding some short descriptions.  we could throw out a lot of words from the paper that way}}
\label{fig:designspace}
\end{figure*}

\subsection{Outputs}

We describe potential pipe-based outputs organized by the five senses: visual, aural, haptic, and olfactory/gustatory.  In addition to outputs we mention explicitly, any outputs possible using traditional electronic components are possible as described above.

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/placeholder/direct.jpg}
\caption{By attaching a servo to a gate structure, it is possible to direct fluid through that pipe.  The gate not only has binary positions (a) and (b), but can direct fractional flow, as well (c).  This could be used to create logic gates, as in (d).  These could be used, for example, in science exhibits.}
\label{fig:direct}
\end{figure}

\emph{Visual} outputs can be mediated by gases, liquids, solid, particulates, or threadables.  Pneumatic actuation (as in \cite{Yao-pneui} or \cite{Harrison-buttons}) provides output, but gases carrying smoke (as in \cite{Sodhi-aireal} demo) also offers visual feedback.  Splitting and mixing of colored liquids can be used as a display tactic; additionally mechanical motion can be fluid-mediated.  Liquids can also be mixed in specific concentrations by servo-controlled gate structures as in Figure \ref{fig:direct}.  Printed optics \cite{Willis-printedoptics} explored visual output at the ends of pipes, and   EL wire can be threaded through custom paths to create neon signs, as in Figure \ref{fig:neon}.

\emph{Aural} outputs can be created with gases.  Helmholz resonance can be leveraged to create computer-controllable sound chambers, as in Zoran's flute \cite{Zoran-flute}.  All that is necessary for this is an enclosed chamber with a pipe connected to it via a narrow neck-point (see Figure \ref{fig:ocarina}).  Passive sound amplification and sound redirection are also possible using pipes.

\begin{figure}[h]
\centering
    \includegraphics[width=3.4in]{figures/placeholder/helmholz.jpg}
\caption{This box (a) contains an enclosed Helmholz resonator, seen in the cutaway view (b).  This resonator creates a sound when air is blown through the pipe at the top, because air compresses at the small neck joint between the chamber and the pipe.  \valkyrie{need to get a test print of this that actually works}}
\label{fig:ocarina}
\end{figure}

\emph{Haptic} outputs are possible through compressible and incompressible fluids.  Fluids can actuate particulates and create programmable pliability \cite{Follmer-jamming}.  Additionally, use of semi-closed pipes with rubberlike material as the caps allows for gases to actuate surface features (see Figure \ref{fig:breathe}).  Free-air haptic feedback \cite{Sodhi-aireal} and volumetric interactions \cite{Iwata-volflex} can be pipe-mediated.

\emph{Olfactory} and \emph{gustatory} outputs can be created through the mixing and splitting of pipes carrying scented or flavored fluids, respectively.

\subsection{Inputs}

Pipes create opportunities for many types of input sensing across the surface of printed devices.  We describe sensing touch, pressure, grasp, flexing, tapping, and manipulation of traditional electronic components.  There are likely more input sensing techniques, both existing and on the horizon, that are compatible with the use of pipes.

\emph{Touch} sensing can be enabled through pipes filled with conductive media.  Traditional wires (threadable) may be used, but conductive paint (liquid) may be more flexible.  Using conductive paint and Swept Frequency Capacitive Sensing (SFCS \cite{Sato-touche}), an interior star pipe topology can enable single-wire touch- and grasp-sensing at any set of points on a printed object's surface (see Figure \cite{fig:toys}).  Simple capacitance measurements are possible using the same media, and custom-designed sensors like those created in Savage, et al.,'s Midas system are also possible \cite{Savage-midas}.

\emph{Pressure} can be sensed in multiple ways.  Slyper, et al.,'s primitives can be attached as the terminus of any semi-closed pipe; the user can manipulate the printed endpoint and the system can sense air pressure changes \cite{Slyper-pressure}.  Pressure can also be sensed via capacitance.

\emph{Grasp} sensing can be enabled using Wimmer's FlyEye techniques \cite{Wimmer-flyeye} or those described in Jamming User Interfaces \cite{Follmer-jamming}.  The optical links necessary for sensing can be created either via clear solid cores as in \cite{Willis-printedoptics} or via fiber optic cables threaded through hollow pipes (see Figure \ref{fig:pens}).

Some 3D printers can fabricate rubber-like materials.  Similar to Slyper, et al., in \cite{Slyper-shape}, we can sense \emph{flexing} and \emph{bending} of prints made on these machines: the ability to 3D print devices to sensing bending and flexing significantly cuts molding and assembly time.

\emph{Tapping} is another input possibility.  Due to differential sound conductivity in different printed materials, pipes of a greater conducting material can be embedded in a model of a lesser conducting material.  A microphone or piezo placed at the system-side of these sound-conducting pipes can determine where the model was tapped and how it was tapped.  Active acoustic sensing as described in Touch \& Activate \cite{Ono-touchandactivate} is also possible; this technique would allow for particular areas of interest (connected to sound-conducting pipes) to be touch sensitive, while other areas (made of non-conducting material) are not sensed.

Sensing via traditional electronic components (e.g., potentiometers, alcohol gas sensors) can be accomplished by conductive inserted media.  Components can be recessed into a print's surface, with their leads implanted into open pipes.  By inserting liquid copper paint instead of threading traditional wires, all components in a model can share a ground line, and the paint's drying process obviates the use of solder or glue to affix the components in place (see Figure \ref{fig:radio}).

\subsection{Fabrication Techniques}

\bjoern{The biggest question wrt fabrication for me is - how do I print these cavities? So general support material avoidance or removal strategies should be more prominent.

(I think the lye bath would probably work better than the high pressure water for our objet prints)}
The physical fabrication process used by different machines gives rise to differing design guidelines.  The two machines we used to create our example objects were the Objet Connex 260 and the Makerbot Replicator 2.

The Objet is an inkjet-based 3D printer which jets drops of polymer in very fine layers, curing each layer with an ultraviolet lamp.  It offers a wide range of material types, including clear/opaque, flexible/hard, and black/white.  Additionally, it permits the mixing of two different materials in a single print: a mixed material could include 70\% flexible and 30\% hard, for example, to be firm but somewhat flexible.  The flexible materials can be inflated when they are thin: they elongate 200\% before breaking.  We have found $.05$inches to be a good thickness for inflatable materials.  Although this machine has many extraordinary capabilities, it unfortunately cannot print overhangs of greater than $14^{\circ}$ without support material.  The support structure used by this machine is a loose matrix of material that can be removed by hand from the surface or blasted out with a water jet from the interior.  The support is also soluble in lye.

The Makerbot Replicator 2 is a fused-deposition modeling (FDM) printer, in which we use exclusively Polylactic Acid (PLA) bioplastics.  More material properties are being explored in PLA: there are now transparent and flexible materials.  They cannot be blended like those in the Objet.  The Makerbot has a fairly coarse resolution compared to the Objet, however it permits overhangs up to $45^{\circ}$ without support and also allows bridging (short spans of material that bridge the gap between two structures with no support underneath).  Due to the FDM process it can even recover from greater overhangs and longer bridges, as the material builds up and can reorganize into a structure \valkyrie{how do I say this?  do I need a photo?  print recovery is pretty awesome-looking.}.  The support used by the Makerbot is made of the same material as the model itself, just printed in a different pattern and attached over as small a surface area as possible to ease removal.  Support inside of a pipe may require pliers, or be infeasible to remove dependent upon geometry.

Assembly is easiest on both machines when the model does not require support material.  Models can be ``cut up'' \bjoern{This negates some of the benefits compared to other manufacturing techniques.} into assembleable pieces to evade support requirements or ease material removal.  We believe that an automatic tool for this task would be highly beneficial, however it is outside the scope of this work.

\subsection{Identity}
Pipes included in objects can have additional effects beyond their designed intentions.

@@

### 2014-03-27 paper/introduction.tex
paper workin

@@

\begin{itemize}
\item We lay out a design space of pipes and their openings, contained media, topologies, and design focus.
\item We offer algorithms and techniques for routing pipes, as well as a design tool implementing them.
\item We describe the interaction possibilities of pipes, including prior work as well as uncharted possibilities.
\item We showcase a set of examples, enabled by our modeling tool and exploring new points in the design space.
\end{itemize}

>>>>

Many existing sensing and actuation approaches, such as FlyEye \cite{Wimmer-flyeye}, Jamming User Interfaces \cite{Follmer-jamming}, and Touch\'{e} \cite{Sato-touche}, can be enhanced by the use of pipes.  Pipes allow makers utilizing these sensing and actuation strategies to locate input and output at arbitrary points on an object's surface.  Thus, we leverage these existing techniques for sensing and actuation while our work's novelty is in \emph{internal modeling} for 3D printing and the \emph{redirection} of IO to arbitrary locations on a device.

@@

\begin{itemize}
\item We lay out a design space of pipes and their openings, contained media, topologies, and design focus.
\item We offer algorithms and techniques for routing pipes, as well as a design tool implementing them.
\item We showcase a set of examples, enabled by our modeling tool and exploring new points in the design space.
\item We describe the interaction possibilities of pipes, including prior work as well as uncharted possibilities.
\end{itemize}

### 2014-03-27 paper/relatedwork.tex
paper workin

@@

Previous work has investigated the integration of interactivity with the fabrication process.  Printed Optics (\cite{Willis-printedoptics}) used light pipes, solid clear tubes, to create integrated displays: our work is inspired by this and goes towards a design tool for these capabilities \valkyrie{bad phrasing; don't de-fang the tool at the front of the paper!}; Printed Optics is a part of our design space, which includes other uses of pipes.  Willis, et al., also did not offer a design tool or techniques for creating light pipes automatically.  Sauron (\cite{Savage-sauron}) investigated the use of computer vision to track interactions with physical mechanisms.  While Savage, et al., used mirrors for redirection, we believe that the required line-of-sight operation for their computer vision techniques may be possible using printed optics or inserted optical fibers.  Slyper, et al., \cite{Slyper-pressure} 3D printed flexible robot armatures capable of sensing interaction via changes in air pressure.  This technique is complementary to ours, as we allow for placing these interactive elements anywhere on an object's surface while they had to place them in locations accessible for electronics insertion.

@@

General-purpose routing for connecting 2D capacitive sensors to terminals is discussed in Midas (\cite{Savage-midas}) \bjoern{Maybe move discussion of Midas to prior section since it really is about fabricating interactive devices. And then here just quickly review that lots of work has gone into routing algorithms for PCBs but that they assume multiple stacked planes connected with vertical vias, not general 3d routing.}.  These techniques, based on circuitboard routing, are limited to planar and unfoldable shapes, and are intended for use on the surface of objects rather than through their cores.

Similar to our neon-sign routing algorithm, Wong presents a graph-based approach for generating continuous line illustrations from images \cite{Wong-continuousline}. Its approach of linking edges into a continuous line based on semi-Eulerization is similar to ours.  However, we are not confined to the plane to avoid line-crossings, and we have additional freedom in creating invisible lines where none existed.  This is possible for us thanks to \emph{blocking out}\footnote{In neon sign making, segments of a tube that should not appear illuminated are covered with opaque paint. This is called "blocking out". \url{http://en.wikipedia.org/wiki/Neon_sign}}. Prior work used tours of nodes, rather than edges, and cast continuous line drawing as a traveling salesman problem~\cite{Bosch-tsp}.  Schweikardt, et al., also made a connection between electroluminescent wire and graph theory, though their contribution is orthogonal to ours: they provide a tangible construction kit with smart nodes for building physical graph examples in which edges are lit by EL wire \cite{Schweikardt-tei09}.

Prior work on modeling the interior of objects (as opposed to their exterior) has been limited.  \valkyrie{ryan, can you give citations here?  I know there is that one about balancing, but Tovi mentioned a ``fruit slicing" one?}  In Make it Stand, material is removed from the interior of objects with the goal of changing their center of gravity \cite{Prevost-makeitstand}.  Bickel, et al., explored enforcing deformation behavior with material blending \cite{Bickel-deformation}, while B\"{a}cher, et al., built an algorithm for automatically placing 3D fabricatable joints into meshes \cite{Bacher-articulated}.  Sauron automatically modified the interior of existing mechanical objects to make them amenable to vision-based sensing \cite{Savage-sauron}, and Infrastructs created encoded interior cavities in 3D printed objects designed to be sensed with a terahertz imaging tool.  Our work differs in that we want to allow users to design pipes and integrate interactive objects with them.

\subsection{Misc}
\bjoern{These may not belong in related work, but I'm just collecting them here.}
\begin{itemize}
\item \cite{strattman1997neon} - Strattman is the standard "text book" for neon lighting designers.
\end{itemize}

>>>>

\valkyrie{can you offer suggestions?  I can imagine the latex buttons paper, touche, acoustic barcodes, capacitive fingerprinting \cite{Harrison-capfingerprint}, printed optics for harrison/poupyrev, and for baudisch maybe.... TDR \cite{Wimmer-TDR}? fiberio \cite{Holz-fiberio}? fabrickation? I'm not sure on this.}

@@

Previous work has investigated the integration of interactivity with the fabrication process.  Printed Optics (\cite{Willis-printedoptics}) used light pipes, solid clear tubes, to create integrated displays: our work is inspired by this and goes towards a design tool for these capabilities; Printed Optics is a part of our design space, which includes other uses of pipes.  Willis, et al., also did not offer a design tool or techniques for creating light pipes automatically.  Sauron investigated the use of computer vision to track interactions with physical mechanisms \cite{Savage-sauron}.  While Savage, et al., used mirrors for redirection, we believe that the required line-of-sight operation for their computer vision techniques may be possible using printed optics or inserted optical fibers.  Slyper, et al., \cite{Slyper-pressure} 3D printed flexible robot armatures capable of sensing interaction via changes in air pressure.  This technique is complementary to ours, as we allow for placing these interactive elements anywhere on an object's surface while they had to place them in locations accessible for electronics insertion.  General-purpose routing for connecting 2D capacitive sensors to terminals is discussed in Midas \cite{Savage-midas}.  This technique, based on circuitboard routing, are limited to planar and unfoldable shapes, and are intended for use on the surface of objects rather than through their cores.  Acoustic Barcodes explored integrating ridges into digitally fabricated objects, whose acoustic signatures when scraped provided an ID code \cite{Harrison-acoustic}.  Fabrickation works towards rapid fabrication of functional objects by integrating construction kit building blocks \cite{Mueller-fabrickation}.

@@

Circuitboard routing algorithms, such as Lee's maze router algorithm, work only on a 2D surface.  So-called ``3D circuitboards'' are typically 2.5D, in that they comprise a stack of 2D circuitboards connected by vias \valkyrie{citation needed}.

Similar to our neon-sign routing algorithm, Wong presents a graph-based approach for generating continuous line illustrations from images \cite{Wong-continuousline}. Its approach of linking edges into a continuous line based on semi-Eulerization is similar to ours.  However, we are not confined to the plane to avoid line-crossings, and we have additional freedom in creating invisible lines where none existed.  This is possible for us thanks to \emph{blocking out}\footnote{In neon sign making, segments of a tube that should not appear illuminated are covered with opaque paint. This is called "blocking out". \url{http://en.wikipedia.org/wiki/Neon_sign}}. Prior work used tours of nodes, rather than edges, and cast continuous line drawing as a traveling salesman problem~\cite{Bosch-tsp}.  Schweikardt, et al., also made a connection between electroluminescent wire and graph theory, though their contribution is orthogonal to ours: they provide a tangible construction kit with smart nodes for building physical graph examples in which edges are lit by EL wire \cite{Schweikardt-tei09}.  In traditional neon sign making, ``routing'' is a conversation between a graphic designer and a glass bender on what is feasible and comfortable for the latter and what will help him avoid burning his hands, wasting glass, and save bending time \cite{strattman1997neon}.

Prior work on modeling the interior of objects (as opposed to their exterior) has been limited.  In Make it Stand, material is removed from the interior of objects with the goal of changing their center of gravity \cite{Prevost-makeitstand}.  Bickel, et al., explored enforcing deformation behavior with material blending \cite{Bickel-deformation}, while B\"{a}cher, et al., built an algorithm for automatically placing 3D fabricatable joints into meshes \cite{Bacher-articulated}.  Sauron automatically modified the interior of existing mechanical objects to make them amenable to vision-based sensing \cite{Savage-sauron}, and Infrastructs created encoded interior cavities in 3D printed objects designed to be sensed with a terahertz imaging tool.  Our work differs in that we want to allow users to design pipes and integrate interactive objects with them.

### 2014-03-27 paper/tubes.tex
paper workin

To allow makers to design novel objects with pipe-powered interfaces, we created a tool for use on arbitrary 3D models.  This tool allows designers to select exterior connection points of their pipes (see Figure \ref{fig:tool-process-interior}) or to import vector art describing the pipes' interior paths (see Figure \ref{fig:tool-process-exterior}).  Once the maker's selections are made, we create a complete routing using either the exterior connection point method (A*, physical simulation) or the interior path method (graph edge creation, Euler circuit generation).  We thicken our routing to create pipes. \valkyrie{We also apply templates where appropriate (e.g., 3D cross-overs for tubes that intersect in the plane).}  The resultant pipes are subtracted from the original mesh, which can then be 3D printed.

Our tool is implemented as a part of Meshmixer, a consumer 3D mesh editing tool, in C++.

@@

To create pipes in which the location and shape of exterior connection points matters, we first allow makers to select the connection points on the mesh's surface using a brush. Our tool then creates an initial shortest-path routing using A* to estimate the routed distance between points.  This routing is used to create a rod; we run physics-based simulation steps on the rod to minimize its bending energy (and thereby minimize the bend radius of the tubes).

\begin{figure}[h!]
\centering
    \includegraphics[width=3.4in]{figures/placeholder/exterior.jpg}
\caption{An example mesh with exterior connection points selected.  The initial A* routing is drawn in {\color{red}red}, and in {\color{tovi}green} is our physically-based rod after simulation.}
\label{fig:tool-process-exterior}
\end{figure}

@@

Our basic first-pass routing algorithm uses the A* routing algorithm \cite{Hart-Astar}.  The path cost in our implementation is based only on shortest distance between the starting and ending points, without weighting for distance from the surface.  We only permit the routing within the voxel grid of the mesh: i.e., the routing is not permitted to traverse through space outside the mesh itself.

Given this initial routing, we create a rod of slightly longer proportional length.  The rod must be longer, as the A* routing can follow the mesh boundaries and the rod requires extra length to be pushed away from the mesh surface.  While we do not have a theoretical basis for this proportional increase, \valkyrie{yet.  what might it be related to?  we can test how much of the A* path is outside the model, and how much is inside; further we can consider the distance from the A* path that is outside the model to the surface.  I think there's a good way to do this.  It might even just mean iteratively changing the rod length until we get something reasonably short and also smooth.  we can also return to the A* that I wrote which finds a routing through the mesh that respects its boundaries; this would get us closer to what we need for length in the first place.}, we have found 1.15 to be a reasonable multiplier in practice.  A too-long rod is forced to kink and expand further than necessary inside the model, while a too-short rod must be stretched, leading to non-smooth bending.

To allow for easiest insertion of media post-print, our physical simulation step minimizes the bend radius of the generated rods.  In order to force our simulated rods away from the surface of the mesh, we create ``wind'' pushing inward from the exterior of the mesh.  This wind keeps the resulting pipe away from the surface. \valkyrie{Ryan: How does it actually work?}

@@

\subsubsection{Pipes with Non-Round Endpoints}
Our tool does not explicitly allow makers to design the shape of their pipes' endpoints.  However, the larger Meshmixer package does allow brushing arbitrary shapes onto the surface and extruding them inward, as in Figure \ref{fig:meshmixer-endpoint}.

@@

Neon signs, as in Figure \ref{fig:tool-process-interior}(d), require a single long path that passes through many desired edges: for example, in our UIST sign, the path must pass through all the edges that comprise the word ``UIST''.  To allow makers to create novel neon signs, as well as similar objects that require design of the pipe's interior path, we built a tool which leverages graph theory to generate this single long path.

@@

\begin{figure}[h!]
\centering
    \includegraphics[width=3.4in]{figures/interior.pdf}
\caption{An input vector graphics file with the points which cannot be tubed as drawn highlighted in {\color{blue}blue} (a).  The connected graph created by our software (b) and the resulting Euler circuit (c) permit creation of a novel neon sign (d).}
\label{fig:tool-process-interior}
\end{figure}

\subsubsection{Routing}
The interior path routing problem is a version of the Chinese Postman Problem\footnote{The CPP is also known as the route inspection problem: \url{http://en.wikipedia.org/wiki/Route_inspection_problem}}, in which we wish to traverse all edges of the graph described in the user's input with a single Euler circuit, much as a postman needs to walk along every road at least once to deliver mail.  We call our relaxation the Spiderman Postman Problem: we allow the creation of non-existing paths (i.e., the postman may traverse buildings in addition to roads).  If inputted path components are disconnected, we must create edges that connect them; additionally we can create new edges connecting odd-degree vertices rather than simply retracing existing edges.  In the final artifact, all created edges will be blocked out by dark material so the inserted medium is not visible (see Figure \ref{fig:tool-process-interior}).  We offer a short algorithm here, with a more mathematically precise definition and associated proof in the appendix.

@@

In order to connect disconnected subgraphs in the input, we reduce each disconnected subgraph to a point.  We add edges from each subgraph-point to each other subgraph-point that represent the minimum distance between any two points in the expanded subgraphs (where distance is Euclidean distance).  We greedily select the smallest weight edges until all subgraphs are joined into a single graph.  We re-expand the subgraphs.

@@

We believe that a lower total weight matching is possible by connecting components and ensuring edge degree evenness together in a global process, however it this not crucial for our purposes.

Once we have a semi-Eulerian graph, we need to create an Euler tour.  An Euler tour is a path that touches every edge once, beginning and ending at the two nodes of odd degree: in this case, those are the user-selected start and end nodes.  We use a weighted modification of Fleury's algorithm\footnote{\url{http://en.wikipedia.org/wiki/Eulerian_path\#Fleury.27s_algorithm}} for this, where instead of randomly selecting from non-bridge paths at each node, we select the path which turns the least from the most recent path (i.e., we prefer to pass straight through a node, if possible).  This minimizes turns in the final artifact, which eases support material removal and assembly.

@@

\subsubsection{Edges that Intersect in the Plane}
We do not currently redirect edges that intersect in the plane.  Because we have a 3-dimensional canvas in which to work, it would be possible to push overlapping paths into the third dimension, however the angel hair EL wire we have used for our prototypes is so thin as to make this unnecessary.

@@

How we actually cut stuff and make tubes happen. \valkyrie{Ryan?}

>>>>

As mentioned, pipe design is not trivial.  Independent pipes cannot intersect lest their contents be unintentionally mixed.  Pipes' bending energy needs to be minimized to allow for easy insertion of media post-print.  Pipes in general need to avoid the surface of objects to prevent fluid leakage when printed by hobbyist-grade machines.  Finally, for designing pipes that follow a user-specified path, we must ensure certain characteristics of that path (e.g., that the path is connected).

To allow makers to design novel objects with pipe-powered interfaces, we created a tool for use on arbitrary 3D models.  This tool allows designers to select exterior connection points of their pipes (see Figure \ref{fig:tool-process-interior}) or to import vector art describing the pipes' interior paths (see Figure \ref{fig:tool-process-exterior}).  Once the maker's selections are made, we create a complete routing using either the exterior connection point method (A*, physical simulation) or the interior path method (graph edge creation, Euler tour generation).  We thicken our routing to create pipes. \valkyrie{We also apply templates where appropriate (e.g., 3D cross-overs for tubes that intersect in the plane).}  The resultant pipes are subtracted from the original mesh, which can then be 3D printed.

Our tool is implemented as a part of Meshmixer, a consumer 3D mesh editing tool, in C++ \cite{Schmidt-meshmixer}.

@@

To create pipes in which the location and shape of exterior connection points matters, we first allow makers to select the connection points on the mesh's surface. Our tool then creates an initial shortest-path routing using A* to estimate the routed distance between points.  This routing is used to create a rod; we run physics-based simulation steps on the rod to minimize its bending energy (and thereby minimize the bend radius of the tubes).

\begin{figure}[h!]
\centering
    \includegraphics[width=3.4in]{figures/exterior.png}
\caption{An example mesh with exterior connection points selected.  The smoothed A* routing is drawn in {\color{red}red}, and in {\color{tovi}green} is our physically-based rod after simulation.}
\label{fig:tool-process-exterior}
\end{figure}

@@

Our basic first-pass routing algorithm uses the A* routing algorithm \cite{Hart-Astar}.  This algorithm keeps a priority queue of points to check for routing based on a cost function; for each point it tags the distance from the start (e.g., startPoint = 0, startNeighbors=1, ...).  Once the end point is reached, we backtrack through the tagged points, always moving towards a lower-valued point.  The path cost in our implementation is based only on shortest distance between the starting and ending points, without weighting for distance from the surface, however the routed path is forced to stay inside the mesh's boundaries.  We route on the voxelized 3D grid of the mesh, and we use a resolution of 128x128x128 voxels.

Note that pipes, once cut, become ``outside the mesh''.  This means that pipes cut in sequence cannot intersect each other.  However, we do not currently track previously-cut pipes for global optimization: we simply greedily select the best routing per pair of points (i.e., per pipe).

Given our initial routing, we create a rod of the same length.  To allow for easiest insertion of media post-print, our physical simulation step minimizes the bend radius of the generated rod.  In order to force our simulated rod away from the surface of the mesh, we create ``wind'' pushing inward from the exterior of the mesh.  This wind keeps the resulting pipe away from the surface. \valkyrie{Ryan: How does it actually work?}

@@

\subsubsection{Pipes with Non-Round Endpoints}
Although start and end radius of pipes is configurable, our tool does not explicitly allow makers to design the \emph{shape} of their pipes' endpoints.  However, the larger Meshmixer package does allow brushing arbitrary shapes onto the surface and extruding them inward, as in Figure \ref{fig:meshmixer-endpoint}.  This technique can be used to create specially-shaped touch pads or to recess the footprint of an electronic component for it to sit flush with the object's surface.

@@

Neon signs, as in Figure \ref{fig:tool-process-interior}(d), require a single long path that passes through many desired edges: for example, in our UIST sign, the path must pass through all the edges that comprise the word ``UIST''.  To allow makers to create novel neon signs, as well as similar objects that require design of the pipe's interior path (for example the maze in Figure \ref{fig:maze}), we built a tool which leverages graph theory to generate this single long path.

@@

\begin{figure}[h!]
\centering
    \includegraphics[width=3.4in]{figures/interior.pdf}
\caption{An input vector graphics file with the points which cannot be tubed as drawn highlighted in {\color{blue}blue} and the start and end points highlighted in {\color{gray}gray} (a).  The connected graph created by our software (b) and the resulting Euler circuit (c) permit creation of a novel neon sign (d).}
\label{fig:tool-process-interior}
\end{figure}

\subsubsection{Routing}
In graph theory, a semi-Eulerian graph is one which has exactly two edges of odd degree.  This kind of graph supports an Euler tour: a path that touches every edge exactly once, beginning and ending at the two nodes of odd degree (in this case, the user-selected start and end nodes).  To allow for a single path through a pipe for post-print media insertion, we need to find a semi-Eulerian graph for which a user's input graphics are a subset, and subsequently find an Euler tour on that graph.

The interior path routing problem is a version of the Chinese Postman Problem\footnote{The CPP is also known as the route inspection problem: \url{http://en.wikipedia.org/wiki/Route_inspection_problem}}, in which we wish to traverse all edges of the graph described in the user's input with a single Euler circuit, much as a postman needs to walk along every road at least once to deliver mail.  We call our relaxation the Spiderman Postman Problem: we allow the creation of non-existing paths (i.e., the postman may traverse buildings in addition to roads).  If input path components are disconnected, we must create edges that connect them; additionally we can create new edges connecting odd-degree vertices rather than simply retracing existing edges.  In the final artifact, all created edges will be blocked out by dark material so the inserted medium is not visible (see Figure \ref{fig:tool-process-interior}).  We offer a short algorithm here, with a more mathematically precise definition and associated proof in the appendix.

@@

In order to connect disconnected subgraphs in the input, each disconnected subgraph's vertices and edges are contracted to a single vertex.  Each of these subgraph-vertices has no outgoing edges, as the expanded subgraphs are disconnected.  We then add one edge from each subgraph-vertex to each other subgraph-vertex, whose weight is the minimum distance between any two vertices in the expanded subgraphs (where distance is Euclidean distance).  We greedily select the smallest weight edges until all subgraphs are joined into a single graph.  We re-expand the subgraphs.

@@

We believe that a lower total weight matching is possible by connecting components and ensuring edge degree evenness together in a global process, as well as by using minimum-weight matching rather than greedy selection, however this optimization is not crucial for our purposes.

Once we have a semi-Eulerian graph, we need to create an Euler tour.  We use a weighted modification of Fleury's algorithm for finding Euler tours\footnote{\url{http://en.wikipedia.org/wiki/Eulerian_path\#Fleury.27s_algorithm}} for this.  In the classic algorithm, each selected edge is removed to create a reduced graph.  To select the next edge while at a node, the algorithm randomly chooses from all incident non-bridge edges (where a bridge edge would create disconnected non-trivial subgraphs if it were removed).  This ensures that every edge is both reachable and traversed.  In our modification, instead of randomly selecting from non-bridge paths at each node, we select the non-bridge edge which turns the least from the most recent path (i.e., we prefer to pass straight through a node, if possible).  This minimizes turns in the final artifact, which eases support material removal and assembly.

@@

\subsubsection{Edges that Intersect in the Plane}
We do not currently redirect edges that intersect in the plane.  Because we have a 3-dimensional canvas in which to work, it would be possible to push overlapping paths into the third dimension.  However, we leave this to future work: for our specific applications so far, it has been unnecessary.

@@

Given either our rod-based or image-based routing, we create cylinders which use the routings as a sweep path.  At each point in the routing, we create a cylinder tall enough to reach to the next point which is oriented along the line segment connecting its origin point to its destination point.  This cylinder's radius is set by the user.  For the rod-based routing, we allow independent start and end radii; in this case, the cylinder's radius is $\frac{k}{n}*\frac{r_1}{r_2}$, where $k$ is the point's index in the routing, $n$ is the total number of points in the routing, $r_1$ is the starting radius, and $r_2$ is the ending radius.

All cylinders generated are joined into a large rod.  This rod is voxelized, along with the original mesh.  These two voxel grids are subtracted, and the result is re-meshed and smoothed.
